Exploring Rural Shrink Smart Through Guided Discovery Dashboards

### 2022-07-31 SDSS Journal Paper/Paper Draft.Rmd
Conclusion stuff

\section {Conclusions}
Specific takeaways - visualization research (broad) vs. dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process. 
\db{In conclusion, }

@@

\subsection{User Testing}
In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices. We examine which aspects of dashboard design ``common knowledge'' hold up in practice, and assess why these principles may not hold in our use-case.

Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.

Using the framework from "Creating design guidelines for building city dashboards from a user's perspectives" [@lee], we will help summarize a users concerns utilizing six main critiques, which are framed into questions. We will adapt these questions to better help the town leaders communicate clear feedback for our team to use. We will collect how the users are experiencing the app with these set questions:
\begin{itemize}
\item Is the information the dashboard presents clear? \textit{(Scope)}
\item Is the dashboard easy to navigate? \textit{(Usability)}
\item Is the dashboard used in a socially responsible manner? \textit{(Ethics)}
\item Do you believe that you can trust the Rural Shrink Smart dashboard? \textit{(Validity)}
\item Do you find value in the Rural Shrink Smart dashboard? \textit{(Utility)}
\end{itemize}

\subsubsection{Beta Testing}
Our Beta testing of a design dashboard before deploying the product to our town leaders, we allowed users from the following groups:
\begin{itemize}
    \item Graphics Group Members (Users know Statistical Graphics)
    \item Rural Shrink Smart Project Members (Users know the project with little background in Statistical Graphics)
    \item  Others (Member of doctoral community that has background in educational adaptivity)
\end{itemize}

While this may not be a large sample size, our range of testers helped with changing approaches in data collection and dashboard changes.

\section{Discussion and Future Research}

\subsection{Data Collection Revisited}

As mentioned above, we wanted to use Unsupervised Clustering Methods to determine similar towns, however we were able to use this methodology to determine data quality in our open data sources. This work empowered our time to look for other open data sources to validate our vital statistics and maps. Since the clustering methods were used differently, we decided to use the USDA's version of rural area definitions. This allowed for a more transparent look into towns that may be similar based on there rurality index. Similar to the Census, the rurality metrics are county based and updated every ten years. This could later cause a problem for towns that may be shrinking faster than the collection of data. It could be worth looking into developing a proxy from our data collected that would be reliable for the years in between the USDA's collected years.

In a related idea of making sure that the towns are identified correctly, we should look into the aging of our towns. This may be important to identify the reasons for shrinking. In the event, that the towns are aging, will we be able to identify how fast the younger generation is moving out of the towns. This could be the most important piece of truly identifying the use of data in towns with an aging population.

\subsection{Visualization Revisited}
In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. As a result, we decided to look for easier ways to display data, such as boxplots and barplots. This direction will be limited but may be useful for the sake of ease on the town leaders.

In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.

\subsection{Feedback}
We will continue to engage with town leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.

>>>>

\section {Conclusions}
Specific takeaways - visualization research (broad) vs. dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.

\db{In conclusion, using a less known statistical graphic, our ideas of using Guided Dscovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. As a result, we decided to look for easier ways to display data, such as boxplots and barplots. This direction will be limited but may be useful for the sake of ease on the town leaders.}

\db{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.}

\db{In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.}

\db{With the lack of research in dashboard design, it could be challenging to develop a dashbaord with a clear layout and clear intention. This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful. This is a generalization not only for our research but along all types of dashboard design. We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.}

@@

\db{When showing our apps to internal research members, it was difficult to collective incoporate what was wanted and what was needed to empower collaborators to help brainstorm. Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout. At times, collaborators could feel as if there suggestions were not being heard or dismissed. Along with the intention of the new designs could have been lost in translation due to the lack of being incoporated earlier or more often. As a result both groups become fustrated.  }

<!-- \subsection{User Testing} -->
<!-- In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices. We examine which aspects of dashboard design ``common knowledge'' hold up in practice, and assess why these principles may not hold in our use-case.  -->

<!-- Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations. -->

<!-- Using the framework from "Creating design guidelines for building city dashboards from a user's perspectives" [@lee], we will help summarize a users concerns utilizing six main critiques, which are framed into questions. We will adapt these questions to better help the town leaders communicate clear feedback for our team to use. We will collect how the users are experiencing the app with these set questions: -->
<!-- \begin{itemize} -->
<!-- \item Is the information the dashboard presents clear? \textit{(Scope)} -->
<!-- \item Is the dashboard easy to navigate? \textit{(Usability)} -->
<!-- \item Is the dashboard used in a socially responsible manner? \textit{(Ethics)} -->
<!-- \item Do you believe that you can trust the Rural Shrink Smart dashboard? \textit{(Validity)} -->
<!-- \item Do you find value in the Rural Shrink Smart dashboard? \textit{(Utility)} -->
<!-- \end{itemize} -->

<!-- \subsubsection{Beta Testing} -->
<!-- Our Beta testing of a design dashboard before deploying the product to our town leaders, we allowed users from the following groups: -->
<!-- \begin{itemize} -->
<!--     \item Graphics Group Members (Users know Statistical Graphics) -->
<!--     \item Rural Shrink Smart Project Members (Users know the project with little background in Statistical Graphics) -->
<!--     \item  Others (Member of doctoral community that has background in educational adaptivity) -->
<!-- \end{itemize} -->

<!-- While this may not be a large sample size, our range of testers helped with changing approaches in data collection and dashboard changes. -->

<!-- \section{Discussion and Future Research} -->

<!-- \subsection{Data Collection Revisited} -->

<!-- As mentioned above, we wanted to use Unsupervised Clustering Methods to determine similar towns, however we were able to use this methodology to determine data quality in our open data sources. This work empowered our time to look for other open data sources to validate our vital statistics and maps. Since the clustering methods were used differently, we decided to use the USDA's version of rural area definitions. This allowed for a more transparent look into towns that may be similar based on there rurality index. Similar to the Census, the rurality metrics are county based and updated every ten years. This could later cause a problem for towns that may be shrinking faster than the collection of data. It could be worth looking into developing a proxy from our data collected that would be reliable for the years in between the USDA's collected years. -->

<!-- In a related idea of making sure that the towns are identified correctly, we should look into the aging of our towns. This may be important to identify the reasons for shrinking. In the event, that the towns are aging, will we be able to identify how fast the younger generation is moving out of the towns. This could be the most important piece of truly identifying the use of data in towns with an aging population. -->

<!-- \subsection{Visualization Revisited} -->
<!-- In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. As a result, we decided to look for easier ways to display data, such as boxplots and barplots. This direction will be limited but may be useful for the sake of ease on the town leaders. -->

<!-- In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey. -->

<!-- \subsection{Feedback} -->
<!-- We will continue to engage with town leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas. -->

### 2022-07-31 SDSS Journal Paper/Paper-Draft.tex
Conclusion stuff

Specific takeaways - visualization research (broad) vs.~dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.
{\textcolor{blue}{In conclusion, }}

@@

\subsection{User Testing}

In this paper, we present the results from testing this application with users, and examine the implications for the feedback we received on our application design and on the more general design of dashboards for statistical novices. We examine which aspects of dashboard design ``common knowledge'\,' hold up in practice, and assess why these principles may not hold in our use-case.

Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'\,' design principles stemming from more controlled experiments hold up in real-world situations.

Using the framework from ``Creating design guidelines for building city dashboards from a user's perspectives'' \citep{lee}, we will help summarize a users concerns utilizing six main critiques, which are framed into questions. We will adapt these questions to better help the town leaders communicate clear feedback for our team to use. We will collect how the users are experiencing the app with these set questions:

\begin{itemize}
\item Is the information the dashboard presents clear? \textit{(Scope)}
\item Is the dashboard easy to navigate? \textit{(Usability)}
\item Is the dashboard used in a socially responsible manner? \textit{(Ethics)}
\item Do you believe that you can trust the Rural Shrink Smart dashboard? \textit{(Validity)}
\item Do you find value in the Rural Shrink Smart dashboard? \textit{(Utility)}
\end{itemize}

\subsubsection{Beta Testing}

Our Beta testing of a design dashboard before deploying the product to our town leaders, we allowed users from the following groups:

\begin{itemize}
    \item Graphics Group Members (Users know Statistical Graphics)
    \item Rural Shrink Smart Project Members (Users know the project with little background in Statistical Graphics)
    \item  Others (Member of doctoral community that has background in educational adaptivity)
\end{itemize}

While this may not be a large sample size, our range of testers helped with changing approaches in data collection and dashboard changes.

\section{Discussion and Future Research}

\subsection{Data Collection Revisited}

As mentioned above, we wanted to use Unsupervised Clustering Methods to determine similar towns, however we were able to use this methodology to determine data quality in our open data sources. This work empowered our time to look for other open data sources to validate our vital statistics and maps. Since the clustering methods were used differently, we decided to use the USDA's version of rural area definitions. This allowed for a more transparent look into towns that may be similar based on there rurality index. Similar to the Census, the rurality metrics are county based and updated every ten years. This could later cause a problem for towns that may be shrinking faster than the collection of data. It could be worth looking into developing a proxy from our data collected that would be reliable for the years in between the USDA's collected years.

In a related idea of making sure that the towns are identified correctly, we should look into the aging of our towns. This may be important to identify the reasons for shrinking. In the event, that the towns are aging, will we be able to identify how fast the younger generation is moving out of the towns. This could be the most important piece of truly identifying the use of data in towns with an aging population.

\subsection{Visualization Revisited}

In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. As a result, we decided to look for easier ways to display data, such as boxplots and barplots. This direction will be limited but may be useful for the sake of ease on the town leaders.

In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.

\subsection{Feedback}

We will continue to engage with town leaders to ensure that we are developing a product that will help them make better data-driven decisions that are useful to increasing their quality of life, which may lend itself to help less shrinkage in these rural areas.

>>>>

Specific takeaways - visualization research (broad) vs.~dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.

{\textcolor{blue}{In conclusion, using a less known statistical graphic, our ideas of using Guided Dscovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. As a result, we decided to look for easier ways to display data, such as boxplots and barplots. This direction will be limited but may be useful for the sake of ease on the town leaders.}}

{\textcolor{blue}{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.}}

{\textcolor{blue}{In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.}}

{\textcolor{blue}{With the lack of research in dashboard design, it could be challenging to develop a dashbaord with a clear layout and clear intention. This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful. This is a generalization not only for our research but along all types of dashboard design. We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.}}

@@

{\textcolor{blue}{When showing our apps to internal research members, it was difficult to collective incoporate what was wanted and what was needed to empower collaborators to help brainstorm. Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout. At times, collaborators could feel as if there suggestions were not being heard or dismissed. Along with the intention of the new designs could have been lost in translation due to the lack of being incoporated earlier or more often. As a result both groups become fustrated.  }}
